Structured Versus Decorated Cospans

### 2020-12-16 cospans.tex
changed \oplus to \otimes (for decorations)

\begin{thm}\label{DC}
Let $\A$ be a category with finite colimits and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a_1\to m_1\leftarrow b_1,x_1\in F(m_1))$ and $(a_2\to m_2\leftarrow b_2,x_2\in F(m_2))$ is:
\begin{equation}\label{eq:tensordecoration}
\begin{tikzcd}
& m_1+m_2 & \\
a_1+a_2\ar[ur,"i_1+i_2"] && b_1+b_2,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad x_1\oplus x_2:=\one \xrightarrow{x_1\times x_2}F(m_1)\times F(m_2)\xrightarrow{\phi_{m_1,m_2}}F(m_1+m_2)
\end{equation}
%$M_1$ and $M_2$:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a_1$};
%\node (B) at (1,0) {$m_1$};
%\node (C) at (2,0) {$b_1$};
%\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
%\node (E) at (3,0) {$a_2$};
%\node (F) at (4,0) {$m_2$};
%\node (G) at (5,0) {$b_2$};
%\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i_1$} (B)
%(C) edge node[above]{$o_1$} (B)
%(E) edge node[above]{$i_2$} (F)
%(G) edge node[above]{$o_2$} (F);
%\end{tikzpicture}
%\]
%is the horizontal 1-cell $M_1 \otimes M_2$:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a_1+a_2$};
%\node (B) at (1.5,0) {$m_1+m_2$};
%\node (C) at (3,0) {$b_1+b_2$};
%\node (D) at (4.55,0) {$x_1+x_2 \in F(m_1+m_2)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i_1+i_2$} (B)
%(C) edge node[above]{$o_1+o_2$} (B);
%\end{tikzpicture}
%\]
%\begin{equation}
%x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)
%\end{equation}

@@

\node (E) at (7,0.5) {$(x_1\oplus x_2)\oplus x_3 \in F((m_1{+}m_2){+}m_3)$};
%{$\scriptstyle\phi_{m_1{+}m_2,m_3}(\phi_{m_1,m_2}(x_1,x_2),x_3) \in F((m_1{+}m_2){+}m_3)$};
\node (F) at (7,-0.5) {$x_1\oplus (x_2\oplus x_3)\in F(m_1{+}(m_2{+}m_3))$};
%{$\scriptstyle\phi_{m_1,m_2{+}m_3}(x_1,\phi_{m_2,m_3}(x_2,x_3))  \in F(m_1{+}(m_2{+}m_3))$};

@@

%%% Christina: I earlier dropped notation x+y for decorations because + is already overused and I find it potentially misleading for the reader. I hope the new symbol is OK?
together with the decoration isomorphism between $F(\alpha)((x_1\oplus x_2)\oplus x_3)\stackrel{\cref{eq:tensordecoration}}{=}F(\alpha)(\phi_{m_1{+}m_2,m_3}(\phi_{m_1,m_2}(x_1,x_2),x_3))$ and $x_1\oplus(x_2\oplus x_3)=\phi_{m_1,m_2{+}m_3}(x_1,\phi_{m_2,m_3}(x_2,x_3))$ formed in the same way as the left two-piece part of \cref{eq:decoiso}. 

@@

%\node (E) at (6,0.5) {$x_1\oplus x_2 \in F(m_1+m_2)$};
%\node (F) at (6,-0.5) {$x_2\oplus x_1 \in F(m_2+m_1)$};

@@

\one \ar[ddrr,dashed,"{(x_2\oplus y_2)\odot(x_1\oplus y_1)}"']\ar[r,"x_1{\times}y_1{\times}x_2{\times}y_2"] \& F(m_1){\times}F(n_1){\times}F(m_2){\times}F(n_2)\ar[r,"\phi_{m_1,n_1}\times\phi_{m_2,n_2}"] \& F(m_1{+}n_1){\times}F(m_2{+}n_2) \ar[d,"\phi_{m_1+n_1,m_2+n_2}"] \\

@@

\one \ar[ddrr,dashed,"{(x_2\odot x_1)\oplus(y_2\odot y_1)}"']\ar[r,"x_1{\times}x_2{\times}y_1{\times}y_2"] \& F(m_1){\times}F(m_2){\times}F(n_1){\times}F(n_2)\ar[r,"\phi_{m_1,m_2}\times\phi_{n_1,n_2}"] \& F(m_1{+}m_2){\times}F(n_1{+}n_2) \ar[d, "F\psi\times F\psi"] \\

@@

			\node (c) at (8,0) {$ (x_1\oplus y_1)\odot (x_2\oplus y_2)  \in F((m_1+n_1)+_{b+b'}(m_2+n_2))$};

@@

			\node (c2) at (8,1) {$(y_1\oplus x_1)\odot(y_2\oplus x_2) \in F((n_1+m_1)+_{b'+b}(n_2+m_2))$};

@@

			\node (c3) at (8,2) {$(y_1 \odot y_2) \oplus (x_1 \odot x_2) \in F((n_1+_{b'} n_2)+(m_1+_b m_2))$};

@@

			\node (c5) at (8,-1) {$(x_1 \odot x_2) \oplus (y_1 \odot y_2) \in F((m_1+_b m_2)+(n_1+_{b'} n_2))$};

@@

			\node (c6) at (8,-2) {$(y_1 \odot y_2) \oplus (x_1 \odot x_2) \in F((n_1+_{b'} n_2)+(m_1+_b m_2))$};

@@

\node (A) at (0,0) {$\tau_1 \colon F(\beta +_\beta \beta)((x_1\oplus y_1)\odot(x_2\oplus y_2)) \to (y_1\oplus x_1)\odot(y_2\oplus x_2)$};
\node (B) at (0,-.5) {$\tau_2 \colon F(\hat{\chi})((y_1 \oplus x_1)\odot(y_2 \oplus x_2)) \to (y_1\odot y_2) \oplus (x_1\odot x_2)$};
\node (C) at (0,-1) {$\tau_3 \colon F(\hat{\chi})((x_1 \oplus y_1)\odot(x_2 \oplus y_2)) \to (x_1\odot x_2) \oplus (y_1\odot y_2)$};
\node (D) at (0,-1.5) {$\tau_4 \colon F(\beta)((x_1\odot x_2)\oplus(y_1\odot y_2)) \to (y_1\odot y_2)\oplus(x_1\odot x_2)$};

@@

\scalebox{.8}{$F(\beta\hat{\chi})((x_1 \oplus y_1)\odot(x_2 \oplus y_2))\cong F(\beta)(F\hat{\chi})((x_1 \oplus y_1)\odot(x_2 \oplus y_2))\xrightarrow{F(\beta)(\tau_3)}F(\beta)((x_1\odot x_2) \oplus (y_1\odot y_2))\xrightarrow{\tau_4}(y_1 \odot y_2) \oplus (x_1 \odot x_2)$}

@@

\scalebox{.8}{$F(\hat{\chi}(\beta +_\beta \beta))( (x_1 \oplus y_1)\odot (x_2 \oplus y_2))\cong F(\hat{\chi})F(\beta +_\beta \beta)( (x_1 \oplus y_1)\odot (x_2 \oplus y_2)) \xrightarrow{F(\hat{\chi})(\tau_1)}F(\hat{\chi})(y_1\oplus x_1)\odot(y_2\oplus x_2)\xrightarrow{\tau_2}(y_1 \odot y_2) \oplus (x_1 \odot x_2)$}

@@

\node (A) at (0,0) {$F(\hat{\chi} (\beta \odot \beta))((x_1 \oplus y_1)\odot(x_2 \oplus y_2))$};
\node (B) at (4,0) {$F(\hat{\chi})((y_1 \oplus x_1)\odot(y_2 \oplus x_2))$};
\node (C) at (0,-1) {$F(\beta)((x_1\odot x_2) \oplus (y_1 \odot y_2))$};
\node (D) at (4,-1) {$(y_1\odot y_2) \oplus (x_1\odot x_2)$};

@@

where in the top left corner, we note that $$F(\hat{\chi} (\beta \odot \beta))((x_1 \oplus y_1)\odot(x_2 \oplus y_2)) = F(\beta\hat{\chi})((x_1 \oplus y_1)\odot (x_2 \oplus y_2))$$

>>>>

\begin{thm}\label{DC}
Let $\A$ be a category with finite colimits and let $(F,\phi,\phi_0)\maps (\A,+) \to (\Cat,\times)$ a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ is symmetric monoidal, where the tensor product
\begin{itemize}
\item of two objects $a$ and $b$ is their coproduct $a+b$ in $\A$,
\item of two vertical 1-morphisms $f \maps a \to b$ and $f' \maps a' \to b'$ is $f+f' \maps a+a' \to b+b'$ in $\A$,
\item of two horizontal 1-cells $(a_1\to m_1\leftarrow b_1,x_1\in F(m_1))$ and $(a_2\to m_2\leftarrow b_2,x_2\in F(m_2))$ is:
\begin{equation}\label{eq:tensordecoration}
\begin{tikzcd}
& m_1+m_2 & \\
a_1+a_2\ar[ur,"i_1+i_2"] && b_1+b_2,\ar[ul,"o_1+o_2"']
\end{tikzcd}\qquad x_1\otimes x_2:=\one \xrightarrow{x_1\times x_2}F(m_1)\times F(m_2)\xrightarrow{\phi_{m_1,m_2}}F(m_1+m_2)
\end{equation}
%$M_1$ and $M_2$:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a_1$};
%\node (B) at (1,0) {$m_1$};
%\node (C) at (2,0) {$b_1$};
%\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
%\node (E) at (3,0) {$a_2$};
%\node (F) at (4,0) {$m_2$};
%\node (G) at (5,0) {$b_2$};
%\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i_1$} (B)
%(C) edge node[above]{$o_1$} (B)
%(E) edge node[above]{$i_2$} (F)
%(G) edge node[above]{$o_2$} (F);
%\end{tikzpicture}
%\]
%is the horizontal 1-cell $M_1 \otimes M_2$:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$a_1+a_2$};
%\node (B) at (1.5,0) {$m_1+m_2$};
%\node (C) at (3,0) {$b_1+b_2$};
%\node (D) at (4.55,0) {$x_1+x_2 \in F(m_1+m_2)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$i_1+i_2$} (B)
%(C) edge node[above]{$o_1+o_2$} (B);
%\end{tikzpicture}
%\]
%\begin{equation}
%x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2)
%\end{equation}

@@

\node (E) at (7,0.5) {$(x_1 \otimes x_2) \otimes x_3 \in F((m_1{+}m_2){+}m_3)$};
\node (F) at (7,-0.5) {$x_1 \otimes (x_2 \otimes x_3)\in F(m_1{+}(m_2{+}m_3))$};

@@

together with the decoration isomorphism between $F(\alpha)((x_1 \otimes x_2) \otimes x_3)\stackrel{\cref{eq:tensordecoration}}{=}F(\alpha)(\phi_{m_1{+}m_2,m_3}(\phi_{m_1,m_2}(x_1,x_2),x_3))$ and $x_1 \otimes (x_2 \otimes x_3)=\phi_{m_1,m_2{+}m_3}(x_1,\phi_{m_2,m_3}(x_2,x_3))$ formed in the same way as the left two-piece part of \cref{eq:decoiso}. 

@@

%\node (E) at (6,0.5) {$x_1 \otimes x_2 \in F(m_1+m_2)$};
%\node (F) at (6,-0.5) {$x_2 \otimes x_1 \in F(m_2+m_1)$};

@@

\one \ar[ddrr,dashed,"{(x_2 \otimes y_2)\odot(x_1 \otimes y_1)}"']\ar[r,"x_1{\times}y_1{\times}x_2{\times}y_2"] \& F(m_1){\times}F(n_1){\times}F(m_2){\times}F(n_2)\ar[r,"\phi_{m_1,n_1}\times\phi_{m_2,n_2}"] \& F(m_1{+}n_1){\times}F(m_2{+}n_2) \ar[d,"\phi_{m_1+n_1,m_2+n_2}"] \\

@@

\one \ar[ddrr,dashed,"{(x_2\odot x_1) \otimes (y_2\odot y_1)}"']\ar[r,"x_1{\times}x_2{\times}y_1{\times}y_2"] \& F(m_1){\times}F(m_2){\times}F(n_1){\times}F(n_2)\ar[r,"\phi_{m_1,m_2}\times\phi_{n_1,n_2}"] \& F(m_1{+}m_2){\times}F(n_1{+}n_2) \ar[d, "F\psi\times F\psi"] \\

@@

			\node (c) at (8,0) {$ (x_1 \otimes y_1) \odot (x_2 \otimes y_2)  \in F((m_1+n_1)+_{b+b'}(m_2+n_2))$};

@@

			\node (c2) at (8,1) {$(y_1\otimes x_1) \odot (y_2 \otimes x_2) \in F((n_1+m_1)+_{b'+b}(n_2+m_2))$};

@@

			\node (c3) at (8,2) {$(y_1 \odot y_2) \otimes (x_1 \odot x_2) \in F((n_1+_{b'} n_2)+(m_1+_b m_2))$};

@@

			\node (c5) at (8,-1) {$(x_1 \odot x_2) \otimes (y_1 \odot y_2) \in F((m_1+_b m_2)+(n_1+_{b'} n_2))$};

@@

			\node (c6) at (8,-2) {$(y_1 \odot y_2) \otimes (x_1 \odot x_2) \in F((n_1+_{b'} n_2)+(m_1+_b m_2))$};

@@

\node (A) at (0,0) {$\tau_1 \colon F(\beta +_\beta \beta)((x_1 \otimes y_1)\odot(x_2\otimes y_2)) \to (y_1 \otimes x_1) \odot (y_2 \otimes x_2)$};
\node (B) at (0,-.5) {$\tau_2 \colon F(\hat{\chi})((y_1 \otimes x_1)\odot(y_2 \otimes x_2)) \to (y_1 \odot y_2) \otimes (x_1 \odot x_2)$};
\node (C) at (0,-1) {$\tau_3 \colon F(\hat{\chi})((x_1 \otimes y_1)\odot(x_2 \otimes y_2)) \to (x_1 \odot x_2) \otimes (y_1 \odot y_2)$};
\node (D) at (0,-1.5) {$\tau_4 \colon F(\beta)((x_1\odot x_2)\otimes(y_1\odot y_2)) \to (y_1\odot y_2) \otimes (x_1\odot x_2)$};

@@

\scalebox{.8}{$F(\beta\hat{\chi})((x_1 \otimes y_1)\odot(x_2 \otimes y_2))\cong F(\beta)(F\hat{\chi})((x_1 \otimes y_1)\odot(x_2 \otimes y_2))\xrightarrow{F(\beta)(\tau_3)}F(\beta)((x_1\odot x_2) \otimes (y_1\odot y_2))\xrightarrow{\tau_4}(y_1 \odot y_2) \otimes (x_1 \odot x_2)$}

@@

\scalebox{.8}{$F(\hat{\chi}(\beta +_\beta \beta))( (x_1 \otimes y_1)\odot (x_2 \otimes y_2))\cong F(\hat{\chi})F(\beta +_\beta \beta)( (x_1 \otimes y_1)\odot (x_2 \otimes y_2)) \xrightarrow{F(\hat{\chi})(\tau_1)}F(\hat{\chi})(y_1\otimes x_1)\odot(y_2\otimes x_2)\xrightarrow{\tau_2}(y_1 \odot y_2) \otimes (x_1 \odot x_2)$}

@@

\node (A) at (0,0) {$F(\hat{\chi} (\beta \odot \beta))((x_1 \otimes y_1)\odot(x_2 \otimes y_2))$};
\node (B) at (4,0) {$F(\hat{\chi})((y_1 \otimes x_1)\odot(y_2 \otimes x_2))$};
\node (C) at (0,-1) {$F(\beta)((x_1\odot x_2) \otimes (y_1 \odot y_2))$};
\node (D) at (4,-1) {$(y_1\odot y_2) \otimes (x_1\odot x_2)$};

@@

where in the top left corner, we note that $$F(\hat{\chi} (\beta \odot \beta))((x_1 \otimes y_1)\odot(x_2 \otimes y_2)) = F(\beta\hat{\chi})((x_1 \otimes y_1)\odot (x_2 \otimes y_2))$$
